### 2014-08-14 dep_dat.Rnw
finished tables/figs for simulation assessment after detiding worked

<<>>=

######
# sim conditions to eval
tide_cat <- c('Diurnal', 'Semidiurnal', 'Mixed Semidiurnal')
tide_cat <- factor(tide_cat, levels = tide_cat)
bio_rng <- round(seq(0, 2, length = 3),2)
tide_assoc <- round(seq(0, 2, length = 3), 2)
err_rng_pro <- round(seq(0, 2, length = 3), 2)
err_rng_obs <- round(seq(0, 2, length = 3), 2)

eval_grd <- expand.grid(tide_cat, bio_rng, tide_assoc, err_rng_pro, 
  err_rng_obs)
names(eval_grd) <- c('tide_cat', 'bio_rng', 'tide_assoc', 'err_rng_pro', 
  'err_rng_obs')
save(eval_grd, file = 'eval_grd.RData')

# grid of regression windows to evaluate
dy_wins <- c(1, 3, 8)
hr_wins <- c(6, 12, 24)
td_wins <- c(0.25, 0.5, 1)
wins_grd <- expand.grid(dy_wins, hr_wins, td_wins)
names(wins_grd) <- c('dec_time', 'hour', 'Tide')
save(wins_grd, file = 'wins_grd.RData')

# grid of both eval and window combinations
comb_grd <- expand.grid(tide_cat, bio_rng, tide_assoc, err_rng_pro, err_rng_obs,
  dy_wins,  hr_wins, td_wins)
names(comb_grd) <- c(names(eval_grd), names(wins_grd))
save(comb_grd, file = 'comb_grd.RData')

######
# simulation results

# 'mod_perf' is all results for each simulation/window combination 
# created in detiding_sims.RProj source_me.r
load('mod_perf.RData')

# used in beginning of assessment
perf_means <- colMeans(mod_perf[, c('cor', 'err')])
perf_rngs <- apply(mod_perf[, c('cor', 'err')], 2, range)

@