### 2014-08-22 dep_dat.Rnw
finished tweaks to methods (removal of tidal categories, smaller hour windows), some proofing, added multimedia

<<>>=
######
# simulation results

# 'mod_perf' is all results for each simulation/window combination 
# created in detiding_sims.RProj source_me.r
load('mod_perf.RData')

# used in beginning of assessment
perf_means <- colMeans(mod_perf[, c('cor', 'err')])
perf_rngs <- apply(mod_perf[, c('cor', 'err')], 2, range)

######
# anomalous data, from 'case_res' code chunk in 'swmp_noise.Rnw'
load('met_comp.RData')

anom_obs <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Observed', ]
anom_obs$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_obs$.id)
anom_obs$.id <- factor(anom_obs$.id, levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'), labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))
anom_dtd <- met_comp[met_comp$Metric %in% 'Anom' & met_comp$Input %in% 'Detided', ]
anom_dtd$.id <- gsub('_wtreg_[0-9]*.RData$', '', anom_dtd$.id)
anom_dtd$.id <- factor(anom_dtd$.id, labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 'Sapelo Island'))

######
# objects for selecting window widht combinations for case studies
casewins <- list(6, 1, 0.5)
@