### 2014-08-22 multimedia.tex

The supporting information for this manuscript includes a graphical illustration of the weighting scheme described in the material and procedures section (\href{http://spark.rstudio.com/beckmw/weights_widget}{http://spark.rstudio.com/beckmw/weights\_widget}), results for each simulation (\href{http://spark.rstudio.com/beckmw/detiding_sims}{http://spark.rstudio.com/beckmw/detiding\_sims}), and results for each case study (\href{http://spark.rstudio.com/beckmw/detiding_cases}{http://spark.rstudio.com/beckmw/detiding\_cases}).  Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.