### 2014-08-22 swmp_noise.Rnw

\documentclass[letterpaper,12pt,oneside]{article}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{indentfirst}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{outlines}
\usepackage{lineno}
\usepackage{array}
\usepackage{times}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage[position=t]{subfig}
\usepackage{paralist}
\usepackage[noae]{Sweave}
\usepackage{natbib}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{bm}
\usepackage{showlabels}
\bibpunct{(}{)}{,}{a}{}{,}

% page margins and section title formatting
\linespread{2}
\setlength{\footskip}{0.5in}
\titleformat*{\section}{\Large\bf\em}
\titleformat*{\subsection}{\singlespace\large\bf}
\titleformat*{\subsubsection}{\singlespace\normalsize\bf\em}
\titlespacing{\section}{0in}{0in}{0in}
\titlespacing{\subsection}{0in}{0in}{0in}
\titlespacing{\subsubsection}{0in}{0in}{0in}

% cleveref options
\crefname{table}{Table}{Tables}
\crefname{figure}{Fig.}{Figs.}
\renewcommand{\figurename}{Fig.}

% aliased citations
\defcitealias{BeckIR}{Beck and Hagy, In review}
\defcitealias{CDMO14}{CDMO 2014}
\defcitealias{RDCT14}{RDCT 2014}

% acronyms
\acrodef{DO}[DO]{dissolved oxygen}
\acrodef{NERRS}[NERRS]{National Estuarine Research Reserve System}
\acrodef{RMSE}[RMSE]{root mean square error}
\acrodef{SWMP}[SWMP]{System Wide Monitoring Program}
\acrodef{WRTDS}[WRTDS]{weighted regression on time, discharge, and season}

% assorted functions
% for multiple rows in table headers
\newcommand{\head}[2]{\multicolumn{1}{>{\arraybackslash}p{#1}}{#2}}
% for milligrams per litre
\newcommand{\mgl}{mg L$^{-1}$}

% hides (not removes) numbering for section, subsection, etc.
% left indents
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{}
\renewcommand{\thesubsubsection}{}
\makeatletter
\def\@seccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\sectionignore\@gobbletwo
\def\@subseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\subsectionignore\@gobbletwo
\def\@subsubseccntformat#1{\csname #1ignore\expandafter\endcsname\csname the#1\endcsname\quad}
\let\subsubsectionignore\@gobbletwo
\let\latex@numberline\numberline
\def\numberline#1{\if\relax#1\relax\else\latex@numberline{#1}\fi}
\makeatother

% dependent data
<<dep_dat, eval = T, echo = F, cache = F, results = 'hide', child = 'dep_dat.Rnw'>>=
@

%knitr options
<<setup, cache = F, echo = F>>=
# set global chunk options
opts_chunk$set(fig.align = 'center', message = F, dev = 'pdf', dev.args = list(family = 'serif'), fig.pos = '!ht', warning = F, echo = F)
options(replace.assign=TRUE,width=90,digits=1)
@

\begin{document}

\raggedbottom
% \linenumbers
\raggedright
\urlstyle{same}
\setlength{\parindent}{0.5in}
\renewcommand\refname{References \vspace{12pt}}

%%%%%%
% title page
\input{titlepage.tex}

%%%%%%
% acknowledgments
\section{Acknowledgments}

%%%%%%
% abstract
\centerline{{\bf Abstract}}
\begin{singlespace} \small
\noindent 
\input{abstract.tex}
\normalsize
\end{singlespace}
\noindent {\bf Key words}:

\acresetall
\clearpage

%%%%%%
% intro
\section{Introduction} \label{intro}
<<intro, child = 'intro.Rnw'>>=
@

%%%%%%
% materials and procedures
\section{Materials and Procedures}
<<mat_proc, child = 'mat_proc.Rnw'>>=
@

%%%%%%
% assessment
\section{Assessment}
<<assess, child = 'assess.Rnw'>>=
@

%%%%%%
% discussion
\section{Discussion}
<<discussion, child = 'discussion.Rnw'>>=
@

%%%%%%
% comments and recs
\section{Comments and recommendations}
<<comments, child = 'comments.Rnw'>>=
@

%%%%%%
% refs
\begin{singlespace}
\bibliographystyle{M:/docs/bibtex_bst/apalike_mine}
\bibliography{M:/docs/ref_diss}
\end{singlespace}
\clearpage

%%%%%%
% figures

\section{Figures}

% example of detiding a simulated time series
\centering\vspace*{\fill}
<<do_dtd, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 5, eval.after='fig.cap', fig.cap = 'Example of detiding a simulated \\ac{DO} time series.  Simulated values are those in \\cref{fig:do_sim}.  Yellow indicates daylight periods.'>>=
######
# examples of how detiding works

vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
vec <- seq(vec[1], vec[2], by = 60*30)

# create simulated time series of DO, tide, etc.
DO_sim <- ts_create(
  vec, 
  do.amp = 2, 
  tide_cat = 'Mixed Semidiurnal', 
  tide_assoc = 4,
  err_rng_obs = 2, 
  err_rng_pro = 2, 
  seeded = T
  )  
prd_tmp <- wtreg_fun(DO_sim, wins = list(6, 12, 1))

levs <- c('DO_obs', 'DO_prd', 'DO_nrm')
to.plo <- melt(prd_tmp, id.var = c('Day', 'sunrise'),
  measure.var = levs
  )
to.plo$variable <- factor(to.plo$variable, levels = levs)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p <- ggplot(to.plo, aes(x = Day, y = value, col = sunrise)) +
  geom_line() +
  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
  theme_bw() +
  ylab(ylab) + 
  scale_colour_gradientn(colours = c('orange', 'black')) +
  theme(legend.position = 'none')

facet_wrap_labeller(p, labels = c(
  expression(italic(DO [obs])),
  expression(italic(DO [tid])),
  expression(italic(DO [nrm]))
  ))

@
\vfill
\clearpage

% example of creating simulated time series
\centering\vspace*{\fill}
<<do_sim, eval = T, echo = F, cache = T, fig.width = 8, fig.height = 7, eval.after='fig.cap', fig.cap = 'Example of creating a simulated \\ac{DO} time series.  Values were simulated every 30 minutes for 30 days.  Yellow indicates daylight periods.'>>=
# create time vector
vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
vec <- seq(vec[1], vec[2], by = 60*30)

# create simulated time series of DO, tide, etc.
DO_sim <- ts_create(
  vec, 
  do.amp = 2, 
  tide_cat = 'Mixed Semidiurnal', 
  tide_assoc = 4,
  err_rng_obs = 2, 
  err_rng_pro = 2, 
  seeded = T
  )

levs <- c('e_obs', 'e_pro', 'DO_unc', 'DO_die', 'DO_bio', 'DO_adv', 'DO_obs')
to.plo <- melt(DO_sim, id.var = c('Day', 'sunrise'),
  measure.var = levs
  )
to.plo$variable <- factor(to.plo$variable, levels = levs)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p <- ggplot(to.plo, aes(x = Day, y = value, col = sunrise)) +
  geom_line() +
  facet_wrap(~ variable, scales = 'free_y', ncol = 1) + 
  theme_bw() +
  ylab(ylab) + 
  scale_colour_gradientn(colours = c('orange', 'black')) +
  theme(legend.position = 'none')

facet_wrap_labeller(p, labels = c(
  expression(italic(epsilon [ obs])),
  expression(italic(epsilon [pro])),
  expression(italic(
    paste(DO [unc], '=', epsilon [ obs] + epsilon [ pro]))),
  expression(italic(DO [die])),
  expression(italic(paste(DO [bio], '=', DO [die] + DO [unc]))), 
  expression(italic(DO [adv])),
  expression(italic(paste(DO [obs], '=', DO [bio] + DO [adv])))
  ))
@
\vfill
\clearpage

% plot of representative time series for simulation
\centering\vspace*{\fill}
<<sim_ex, fig.height = 4, fig.width = 7.5, out.width = '\\textwidth', cache = T, echo = F, eval = T, fig.cap = 'Representative examples of simulated observed \\ac{DO} ($DO_{obs}$) and biological \\ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters: tidal strength of tidal association with \\ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.  Parameter values represent the minimum and maximum used in the simulations.'>>=
load('eval_grd.RData')

# find rows in eval_grd of parms to plot
sel_vec <- !with(eval_grd,
  bio_rng %in% 1|
  tide_assoc %in% 1|
  err_rng_pro %in% 1|
  err_rng_obs %in% 1
  )
parms <- eval_grd[sel_vec,]
parms$L1 <- as.numeric(row.names(parms))

# create simulated time series of DO, tide, etc., representatives
vec <- c('2014-05-01 00:00:00', '2014-05-31 00:00:00')
vec <- as.POSIXct(vec, format = '%Y-%m-%d %H:%M:%S')
vec <- seq(vec[1], vec[2], by = 60*30)
to_plo <- alply(parms,
  1,
  .fun = function(x) {
  ts_create(vec, 
    do.amp = x$bio_rng, 
    tide_cat = x$tide_cat, 
    tide_assoc = x$tide_assoc, 
    err_rng_obs = x$err_rng_obs, 
    err_rng_pro = x$err_rng_pro, 
    seeded = T)
    }
  )
names(to_plo) <- rownames(parms)

# prep for plotting
to_plo <- melt(to_plo, id.var = names(to_plo[[1]]))
to_plo <- merge(to_plo, parms, by = 'L1', all.x = T)

# rename extremes for facet labs
labs <- paste('Bio Amp', unique(to_plo$bio_rng))
to_plo$bio_rng <- factor(to_plo$bio_rng, labels = labs)
labs <- paste('Assoc', unique(to_plo$tide_assoc))
to_plo$tide_assoc <- factor(to_plo$tide_assoc, labels = labs)
labs <- paste('Noise pro', unique(to_plo$err_rng_pro))
to_plo$err_rng_pro <- factor(to_plo$err_rng_pro, labels = labs)
labs <- paste('Noise obs', unique(to_plo$err_rng_obs))
to_plo$err_rng_obs <- factor(to_plo$err_rng_obs, labels = labs)

# setup facet labels
facet1_names <- list(
  'Assoc 0' = expression(paste(italic(DO [adv]), ' 0')), 
  'Assoc 2' = expression(paste(italic(DO [adv]), ' 2'))
  )
facet2_names <- list(
  'Bio Amp 0' = expression(paste(italic(DO [die]), ' 0')), 
  'Bio Amp 2' = expression(paste(italic(DO [die]), ' 2'))
  )
facet3_names <- list(
  'Noise pro 0' = expression(paste(italic(epsilon [pro]), ' 0')), 
  'Noise pro 2' = expression(paste(italic(epsilon [pro]), ' 2'))
  )
facet4_names <- list(
  'Noise obs 0' = expression(paste(italic(epsilon [obs]), ' 0')), 
  'Noise obs 2' = expression(paste(italic(epsilon [obs]), ' 2'))
  )
plot_labeller <- function(variable,value){
  if (variable=='tide_assoc')
    return(facet1_names[value])
  if (variable=='bio_rng')
    return(facet2_names[value])
  if (variable=='err_rng_pro')
    return(facet3_names[value])
  if (variable=='err_rng_obs')
    return(facet4_names[value])
  }

cols <- ggplotColours(2)
ggplot(to_plo, aes(x = Day, y = DO_obs, group = L1)) + 
  geom_line(aes(colour = 'Observed'), alpha = 0.7) +
  geom_line(aes(y = DO_bio, colour = 'Biological'), alpha = 0.7) +
  theme_bw() + 
  scale_colour_manual(
    labels = expression(italic(DO [bio]), italic(DO [obs])),
    values = cols
    ) +
  theme(legend.position = 'top', 
    legend.title = element_blank(), 
    axis.text = element_text(size = 8)
    ) + 
  ylab(expression(paste('DO (mg ', L^-1, ')'))) +
  facet_grid(err_rng_obs + err_rng_pro ~ bio_rng + tide_assoc, 
    labeller = plot_labeller) 
@
\vfill
\clearpage

% example of error surfaces 
\centering\vspace*{\fill}
<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying simulation parameters: strength of tidal association with \\ac{DO} signal $DO_{adv}$, amount of process uncertainty $\\epsilon_{pro}$, amount of observation observation uncertainty $\\epsilon_{obs}$, and strength of diel \\ac{DO} component $DO_{die}$.  Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of simulation parameters.  Results are averaged for all window widths used to evaluate the regressions (\\cref{fig:err_surf2}).">>=

# load summary of model performance for simulations
load('mod_perf.RData')

to_plo <- mod_perf

# aggregate for simpler plot
to_plo <- ddply(
  to_plo, 
  .variable = c('bio_rng', 'tide_assoc', 'err_rng_pro', 'err_rng_obs'),
  .fun = function(x) data.frame(cor = mean(x$cor), err = mean(x$err))
  )

# reassign factor labels for bio amp and tidal assoc
labs <- paste('Bio Amp', unique(to_plo$bio_rng))
to_plo$bio_rng <- factor(to_plo$bio_rng, labels = labs)
labs <- paste('Assoc',unique(to_plo$tide_assoc))
to_plo$tide_assoc <- factor(to_plo$tide_assoc, labels = labs)

# setup facet labels
facet1_names <- list(
  'Bio Amp 0' = expression(paste(italic(DO [die]), ' 0')), 
  'Bio Amp 1' = expression(paste(italic(DO [die]), ' 1')),
  'Bio Amp 2' = expression(paste(italic(DO [die]), ' 2'))
  )
facet2_names <- list(
  'Assoc 0' = expression(paste(italic(DO [adv]), ' 0')), 
  'Assoc 1' = expression(paste(italic(DO [adv]), ' 1')),
  'Assoc 2' = expression(paste(italic(DO [adv]), ' 2'))
  )

plot_labeller <- function(variable,value){
  if (variable=='tide_assoc')
    return(facet1_names[value])
  if (variable=='bio_rng')
    return(facet2_names[value])
  }

mat_theme <-   theme(
    panel.margin = unit(0, 'lines'), 
    strip.text = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = 'top',
    axis.text = element_text(size = 8)
    )

p1 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), z = cor, fill = cor)) + 
  geom_tile() +
  facet_grid(bio_rng ~ tide_assoc,
    labeller = plot_labeller) +
  scale_fill_gradientn(name = 'Correlation', 
    colours = brewer.pal(11, 'Spectral'), limits = c(0, 1)
    ) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_discrete(expand = c(0,0)) +
  xlab(expression(italic(epsilon [pro]))) +
  ylab(expression(italic(epsilon [obs]))) + 
  theme_bw() +
  mat_theme

p2 <- ggplot(to_plo, aes(x = factor(err_rng_pro), y = factor(err_rng_obs), z = err, fill = err)) + 
  geom_tile() +
  facet_grid(bio_rng ~ tide_assoc,
    labeller = plot_labeller) +
  scale_fill_gradientn(name = 'RMSE', 
    colours = rev(brewer.pal(11, 'Spectral')), limits = c(0, 2.5)
    ) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_discrete(expand = c(0,0)) +
  xlab(expression(italic(epsilon [pro]))) +
  ylab(expression(italic(epsilon [obs]))) + 
  theme_bw() +
  mat_theme

grid.arrange(p1, p2, ncol = 2)

@
\vfill
\clearpage

% example of error surfaces 
\centering\vspace*{\fill}
<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Correlations and errors (\\ac{RMSE}) for detided \\ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \\ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error between detided and biological \\ac{DO} time series from results for a given combination of window widths.  Results are averaged for all simulation parameters used to evaluate the regressions (\\cref{fig:err_surf2}).">>=

# load summary of model performance for simulations
load('mod_perf.RData')

to_plo <- mod_perf

# aggregate for simpler plot
to_plo <- ddply(
  to_plo, 
  .variable = c('dec_time', 'hour', 'Tide'),
  .fun = function(x) data.frame(cor = mean(x$cor), err = mean(x$err))
  )

# reassign factor labels for windows
labs <- paste('Tide', unique(to_plo$Tide))
to_plo$Tide <- factor(to_plo$Tide, labels = labs)

mat_theme <-   theme(
    panel.margin = unit(0, 'lines'), 
    strip.text = element_text(size = 8),
    legend.text = element_text(size = 8),
    legend.position = 'top',
    axis.text = element_text(size = 8)
    )

p1 <- ggplot(to_plo, aes(x = factor(dec_time), y = factor(hour), z = cor, fill = cor)) + 
  geom_tile() +
  facet_grid(~ Tide) +
  scale_fill_gradientn(name = 'Correlation', 
    colours = brewer.pal(11, 'Spectral')#, limits = c(0.3, 0.7),
#     breaks = seq(0.3, 0.7, by = 0.1)
    ) +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  xlab('Days') +
  ylab('Hours') + 
  theme_bw() +
  mat_theme

p2 <- ggplot(to_plo, aes(x = factor(dec_time), y = factor(hour), z = err, fill = err)) + 
  geom_tile() +
  facet_grid(~ Tide) +
  scale_fill_gradientn(name = 'RMSE', 
    colours = rev(brewer.pal(11, 'Spectral'))#, limits = c(1, 1.5), 
#     labels = format(seq(1, 1.5, by = 0.1), nsmall = 1)
    ) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_discrete(expand = c(0,0)) +
  xlab('Days') +
  ylab('Hours') + 
  theme_bw() +
  mat_theme

grid.arrange(p1, p2, ncol = 2)

@
\vfill
\clearpage

% maps of each case
\centering\vspace*{\fill}
<<case_map, cache = T, fig.height = 5, fig.width = 8, eval = T, echo = F, fig.cap = 'Locations of \\ac{NERRS} sites used as case studies to evaluate of weighted regression.  Individual stations at each reserve are PDBJE (Joe Leary Estuary at Padilla Bay), RKBMB (Middle Blackwater River at Rookery Bay), SAPDC (Dean Creek at Sapelo Island), and TJRBR (Boca Rio at Tijuana River).'>>=

# low do/high tide, high do/low tide, low do/low tide, high do/low tide
cases <- c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC')

# zooms for each case, larger is more zoom
zooms <- c(13, 13, 12, 15)

require(ggmap)

# breaks on map
myBreaks <- function(x){
    breaks <- round(quantile(x, probs = c(0.1, 0.5, 0.9)), 2)
    names(breaks) <- attr(breaks,"labels")
    breaks
  }

for(case in cases){

  case_meta <- get_map_meta(case)

  mapImageData<-get_map(
    location=c(lon=mean(case_meta$Longitude),lat=mean(case_meta$Latitude)),
    source='google',
    maptype='satellite',
    zoom = zooms[case == cases]
    )

  reserve <- trim.trailing(unique(case_meta$Reserve.Name))

  p <- ggmap(mapImageData,
    extent = "device", 
    base_layer=ggplot(data=case_meta, aes(x =Longitude, 
      y=Latitude, group='Station.Code'))
      ) + 
    geom_point(size = 3, colour = 'lightblue') +
    geom_text(data=case_meta,aes(x=Longitude,y=Latitude,label=Station,
      hjust=0.4,vjust=2),size=5, colour = 'lightblue') +
    ggtitle(reserve) +
    theme(axis.title = element_blank(), axis.text = element_text(size = 8)) +
    scale_x_continuous(breaks = myBreaks) +
    scale_y_continuous(breaks = myBreaks)

  assign(paste0('p', which(case == cases)), p)

  }

# for us map
meta <- get_map_meta(cases)
meta <- meta[, c('Longitude', 'Latitude','Reserve.Name')]

# create viewports
# vp.1 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.67, x=0)
# vp.2 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.67, x=0.5)
# vp.3 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.33, x=0)
# vp.4 <- viewport(height=unit(.32, "npc"), width=unit(0.5, "npc"), 
#                            just=c("left","top"), 
#                            y=0.33, x=0.5)
# 
# par(mfrow = c(3, 1), mar = numeric(4))
# grob({
  map('state', col = alpha('black', 0.4), ylim = c(23, 50))
  points(meta[, c('Longitude', 'Latitude')], cex =2, pch = 16)
  text(meta$Longitude, meta$Latitude, meta$Reserve.Name, pos = 4, cex = 1.5)
  map.scale(x = -122, y = 26)
#   })
# print(p1, vp=vp.1)
# print(p2, vp=vp.2)
# print(p3, vp=vp.3)
# print(p4, vp=vp.4)

@
\vfill
\clearpage

% example from SAPHD, phase out
\centering\vspace*{\fill}
<<phase_out, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$). Results are for the Sapelo Island station for a seven day period when high tide events were out of phase with diel periods, creating lower than expected observed \\ac{DO} during night and day periods. Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.'>>=
######
# SAPHD example

load('case_grds.RData')

# subsets by case. date range, window comb
case <- 'SAPDC'
dat.rng<-as.Date(c('2012-02-01','2012-02-14')) 
sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')

load(paste0('wtreg/', sel_vec))
inst_subs <- get(gsub('.RData', '', sel_vec))

tzone <- attr(inst_subs$DateTimeStamp, 'tzone')

inst_subs$Date <- as.Date(inst_subs$DateTimeStamp, tz = tzone)
inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
inst_subs <- inst_subs[inst.rng,]

# category for plotting
inst_subs$cats <- 'out'
inst_subs$cats[inst_subs$Date > as.Date('2012-02-07')] <- 'in'

##
# custom theme, mod of theme_b
my_theme <- theme(
  legend.position = 'top',
  axis.title.x = element_blank(),legend.box= 'horizontal',
  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
  )

# function for setting range on y axis
rng.fun<-function(vec.in){
  rngs<-range(vec.in,na.rm=T)
  buffs<-0.07*abs(diff(rngs))
  c(rngs[1]-buffs,rngs[2]+buffs)
  }

##
# out of phase plots

# DO plot
to_plo <- met.day.fun(inst_subs, case)
to_plo <- to_plo[to_plo$cats %in% 'out', ]
names(to_plo)[names(to_plo) %in% 'variable'] <- 'solar'
ggpoly1 <- poly.fun(to_plo$solar, to_plo, for_leg = T)
ggpoly2 <- poly.fun(to_plo$solar, to_plo)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p1 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
  ggpoly1 +
  geom_line(aes(y = DO_obs, colour = 'Observed')) +
  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
  coord_cartesian(ylim = rng.fun(to_plo$DO_obs)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +
  scale_y_continuous(ylab) +
  my_theme +
  theme(legend.title = element_blank())

#
# tide plot
ylab<-expression(paste('Height (m)'))
p2 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
  ggpoly2 +
  geom_line(aes(y = Tide, colour = TotPAR)) +
  coord_cartesian(ylim = rng.fun(to_plo$Tide)) +
  theme_bw() +  
  my_theme + 
  scale_y_continuous(ylab) +
  scale_colour_gradient("PAR", low='blue', high='yellow',
    guide = guide_colorbar(direction = "horizontal",barheight= 0.4)) +
  theme(legend.title.align = 1)

# get common legend, remove from p1
mylegend1 <- g_legend(p1)
mylegend2 <- g_legend(p2)
p1 <- p1 + theme(legend.position = 'none')
p2 <- p2 + theme(legend.position = 'none')

# Get the widths
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3])

# Set the widths
pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth

# first plot
grid.arrange(
  arrangeGrob(mylegend1, mylegend2, ncol = 1),    
  arrangeGrob(pA, pB, ncol = 1),
  ncol = 1, heights = c(2,5)
)

@
\vfill
\clearpage

% example from SAPHD, phase in
\centering\vspace*{\fill}
<<phase_in, cache = T, fig.height = 3, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Continuous \\ac{DO} time series before (observed) and after (detided) detiding with weighted regression and tidal height colored by total photosynthetically active radiation (mmol m$^{-2}$). Results are for the Sapelo Island station for a seven day period when high tide events were in phase with diel periods, creating higher than expected observed \\ac{DO} during night and day periods. Detided values are based on a weighted regression with half windows of six days, one hour within each day, and tidal height proportion of one half.'>>=
######
# SAPHD example

load('case_grds.RData')

# subsets by case. date range, window comb
case <- 'SAPDC'
dat.rng<-as.Date(c('2012-02-01','2012-02-14')) 
sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')

load(paste0('wtreg/', sel_vec))
inst_subs <- get(gsub('.RData', '', sel_vec))

tzone <- attr(inst_subs$DateTimeStamp, 'tzone')

inst_subs$Date <- as.Date(inst_subs$DateTimeStamp, tz = tzone)
inst.rng <- inst_subs$Date<=dat.rng[2] & inst_subs$Date>=dat.rng[1]
inst_subs <- inst_subs[inst.rng,]

# category for plotting
inst_subs$cats <- 'out'
inst_subs$cats[inst_subs$Date > as.Date('2012-02-07')] <- 'in'

##
# custom theme, mod of theme_b
my_theme <- theme(
  legend.position = 'top',
  axis.title.x = element_blank(),legend.box= 'horizontal',
  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
  )

# function for setting range on y axis
rng.fun<-function(vec.in){
  rngs<-range(vec.in,na.rm=T)
  buffs<-0.07*abs(diff(rngs))
  c(rngs[1]-buffs,rngs[2]+buffs)
  }

##
# out of phase plots

# DO plot
to_plo <- met.day.fun(inst_subs, case)
to_plo <- to_plo[!to_plo$cats %in% 'out', ]
names(to_plo)[names(to_plo) %in% 'variable'] <- 'solar'
ggpoly1 <- poly.fun(to_plo$solar, to_plo, for_leg = T)
ggpoly2 <- poly.fun(to_plo$solar, to_plo)

ylab<-expression(paste('DO (mg ',L^-1,')'))
p1 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
  ggpoly1 +
  geom_line(aes(y = DO_obs, colour = 'Observed')) +
  geom_line(aes(y = DO_nrm, colour = 'Detided')) +
  coord_cartesian(ylim = rng.fun(to_plo$DO_obs)) +
  scale_fill_manual(values='orange',labels='Day') +
  theme_bw() +
  scale_y_continuous(ylab) +
  my_theme +
  theme(legend.title = element_blank())

#
# tide plot
ylab<-expression(paste('Height (m)'))
p2 <- ggplot(to_plo, aes(x = DateTimeStamp)) + 
  ggpoly2 +
  geom_line(aes(y = Tide, colour = TotPAR)) +
  coord_cartesian(ylim = rng.fun(to_plo$Tide)) +
  theme_bw() +  
  my_theme + 
  scale_y_continuous(ylab) +
  scale_colour_gradient("PAR", low='blue', high='yellow',
    guide = guide_colorbar(direction = "horizontal",barheight= 0.4)) +
  theme(legend.title.align = 1)

# get common legend, remove from p1
mylegend1 <- g_legend(p1)
mylegend2 <- g_legend(p2)
p1 <- p1 + theme(legend.position = 'none')
p2 <- p2 + theme(legend.position = 'none')

##
# in phase plots

# Get the widths
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3])

# Set the widths
pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth

# plot
grid.arrange(
  arrangeGrob(mylegend1, mylegend2, ncol = 1),    
  arrangeGrob(pA, pB, ncol = 1),
  ncol = 1, heights = c(2,5)
  )

@
\vfill
\clearpage

% example from SAPDC
\centering\vspace*{\fill}
<<case_ex, cache = T, fig.height = 2.5, fig.width = 5.5, out.width = '0.8\\textwidth', eval = T, echo = F, fig.cap = 'Example of daily mean metabolism (net ecosystem metabolism, gross production, and total respiration) before (observed) and after (detided) detiding with weighted regression. Results are for the Sapelo Island station for a two week period in February, 2012 when high tide was out of phase with the diel cycle during the first week (\\cref{fig:phase_out}) and in phase during the second week (\\cref{fig:phase_in}).'>>=
######
# SAPDC example

load('met_ls.RData')
load('case_grds.RData')

# subsets by case. date range, window comb
case <- 'SAPDC'
dat.rng<-as.Date(c('2012-02-01','2012-02-14')) 
sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
sel_vec <- paste0(case, '_wtreg_', sel_vec, '.RData')

# select case, window
met_subs <- met_ls[[grep(sel_vec, names(met_ls))]]

# subset by date
met.rng <- met_subs$Date<=dat.rng[2] & met_subs$Date>=dat.rng[1]
met_subs <- met_subs[met.rng,]

##
# custom theme, mod of theme_bw

my_theme <- theme(
  legend.title = element_blank(),legend.position = 'top',
  axis.title.x = element_blank(),legend.box= 'horizontal',
  plot.margin= unit(c(0, 1, 0, 1), "lines") # top right bottom left
  )

# function for setting range on y axis
rng.fun<-function(vec.in){
  rngs<-range(vec.in,na.rm=T)
  buffs<-0.07*abs(diff(rngs))
  c(rngs[1]-buffs,rngs[2]+buffs)
  }

##
# metab plot
to_plo1 <- melt(met_subs, id.var = c('Date'), 
  measure.var = grep('Pg|Rt|NEM', names(met_subs), value = T)
  )
to_plo1$Input <- 'Observed'
to_plo1$Input[grep('dtd', to_plo1$variable)] <- 'Detided'
to_plo1$Input <- factor(to_plo1$Input, levels = c('Observed', 'Detided'))
to_plo1$variable <- gsub('_dtd', '', to_plo1$variable)

ylab<-expression(paste('g ',O [2], ' ', m^-2, d^-1))
p1 <- ggplot(to_plo1, aes(x = Date, y = 0.032 * value, group = variable,
    colour = variable)) +
  geom_line() +
  theme_bw() +
  geom_point(size = 2) +
  facet_wrap(~Input, ncol = 1) +
  scale_y_continuous(ylab)  +
  my_theme

print(p1)

@

% plots of summarized metabolism estimates, before/after detiding
% ELKVM, PDBBY
\centering\vspace*{\fill}
<<metab_sum1, cache = T, eval = T, echo = F, fig.height = 5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.  Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and detided \\ac{DO} time series.'>>=

# load metabolism data
load('met_ls.RData')

# subset metab estimates by 30 day window and tide prop 1
load('case_grds.RData')
sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))

met_sub <- met_ls[grep(sel_vec, names(met_ls))]

# creat weekly, monthly, seasonal categories
met_sub <- llply(met_sub, 
  .fun = function(x){

    # add week, month cols
    x$Week <- format(x$Date, '%W')
    x$Month <- format(x$Date, '%m')

    # seas col
    x$Season <- car::recode(x$Month, "c('01','02','03') = 'W'; 
      c('04','05','06') = 'Sp'; 
      c('07','08','09') = 'Su'; 
      c('10','11','12') = 'F'")

    x

  })

# aggregate by weekly, monthly, seasonal categories
# for each case...
agg_sum <- llply(met_sub, 
  .fun = function(x){

    # metabolism column names
    met_cols <- c('Pg', 'Rt', 'Pg_dtd', 'Rt_dtd')

    # melt by weekly, monthly, seasonal cats
    # change here for cats to includes
    x <- melt(x, measure.var = c('Month', 'Season'),
      id.var = c('Date', met_cols)
      )

    agg_res <- dlply(x,
      .var = c('variable'), 
      .fun = function(y){

        out <- vector('list', length = length(met_cols))
        names(out) <- met_cols
        for(met in met_cols){

          tmp <- summarySE(y, measurevar = met, groupvars = 'value',
            narm = T) 
          names(tmp)[names(tmp) %in% met] <- 'mean'
          out[[met]] <- tmp

        }

        # combine metabolism categories 
        out <- melt(out, id.var = names(out[[1]]))
        names(out)[names(out) %in% 'L1'] <- 'var'
        out

      })

    # combine time categories
    agg_res <- melt(agg_res, id.var = names(agg_res[[1]]))
    names(agg_res)[names(agg_res) %in% 'L1'] <- 'cats'

    # separate var column into metabolis, obs/dtd
    agg_res$var[!grepl('dtd',agg_res$var)] <- paste0(
      agg_res$var[!grepl('dtd',agg_res$var)], '_obs')

    agg_res$sub_var <- gsub('^[A-Z,a-z]*_', '', agg_res$var)
    agg_res$var <- gsub('_[a-z]*$', '', agg_res$var)

    agg_res

  })

to_plo <- melt(agg_sum, id.var = names(agg_sum[[1]]))

# reassign factor labels
to_plo$value <- factor(to_plo$value, levels = c('01', '02', '03', '04',
  '05', '06', '07', '08', '09', '10', '11', '12', 'W', 'Sp', 'Su', 'F'), 
  labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12',
    'W', 'Sp', 'Su', 'F')
  )
to_plo$var <- factor(to_plo$var, levels = c('Pg', 'Rt', 'NEM'), 
  labels = c('Pg', 'Rt', 'NEM'))
to_plo$sub_var <- factor(to_plo$sub_var, levels = c('dtd', 'obs'), 
  labels = c('Detided', 'Observed'))
to_plo$L1 <- gsub('_wtreg_[0-9]*.RData$', '', to_plo$L1)
to_plo$L1 <- factor(to_plo$L1, 
  levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'),
  labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 
    'Sapelo Island')
  )

# convert metab data to g m^-2 d^-1
#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
to_plo[, c('mean', 'sd', 'se', 'ci')] <- 0.032 * to_plo[, c('mean', 'sd', 'se', 'ci')]

ylab <- expression(paste('g ', O [2], ' ', m^-2, d^-1))
for(site in 1:2){

  sub_plo <- to_plo[to_plo$L1 %in% levels(to_plo$L1)[site], ]

  p <- ggplot(sub_plo, aes(x = factor(value), y = mean, group = sub_var,
      colour = sub_var)) +
    geom_line() +
    geom_point() + 
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
      width=.3) +
    facet_wrap(cats ~ var, scales = 'free') +
    theme_bw() +
    scale_y_continuous(ylab) + 
    scale_x_discrete(name = element_blank()) +
    theme(
      legend.position = 'top',
      legend.title = element_blank(),
      plot.margin= unit(rep(0, 4), "lines"),
      axis.text = element_text(size = 8)
      ) +
    ggtitle(levels(to_plo$L1)[site])

  assign(paste0('p', site), p)

  }

# get common legend, remove from p1
mylegend <- g_legend(p1)
p1 <- p1 + theme(legend.position = 'none')
p2 <- p2 + theme(legend.position = 'none')

# final plot
grid.arrange(
  mylegend,   
  arrangeGrob(p1, p2, ncol = 1),
  ncol = 1, heights = c(1,10)
  )

@
\vfill
\clearpage

% RKBMB, SAPDC
% plots of summarized metabolism estimates, before/after detiding
\centering\vspace*{\fill}
<<metab_sum2, cache = T, eval = T, echo = F, fig.height =5.5, fig.width = 7, fig.cap = 'Means and standard errors of daily metabolism estimates (gross production, total respiration, net ecosystem metabolism) aggregated by month and season.  Aggregated estimates are for Rookery Bay and Sapelo Island from observed and detided \\ac{DO} time series.  May was removed from Rookery Bay because of incomplete data.'>>=

# load metabolism data
load('met_ls.RData')

# subset metab estimates by 30 day window and tide prop 1
load('case_grds.RData')
sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))

met_sub <- met_ls[grep(sel_vec, names(met_ls))]

# creat weekly, monthly, seasonal categories
met_sub <- llply(met_sub, 
  .fun = function(x){

    # add week, month cols
    x$Week <- format(x$Date, '%W')
    x$Month <- format(x$Date, '%m')

    # seas col
    x$Season <- car::recode(x$Month, "c('01','02','03') = 'W'; 
      c('04','05','06') = 'Sp'; 
      c('07','08','09') = 'Su'; 
      c('10','11','12') = 'F'")

    x

  })

# aggregate by weekly, monthly, seasonal categories
# for each case...
agg_sum <- llply(met_sub, 
  .fun = function(x){

    # metabolism column names
    met_cols <- c('Pg', 'Rt', 'Pg_dtd', 'Rt_dtd')

    # melt by weekly, monthly, seasonal cats
    # change here for cats to includes
    x <- melt(x, measure.var = c('Month', 'Season'),
      id.var = c('Date', met_cols)
      )

    agg_res <- dlply(x,
      .var = c('variable'), 
      .fun = function(y){

        out <- vector('list', length = length(met_cols))
        names(out) <- met_cols
        for(met in met_cols){

          tmp <- summarySE(y, measurevar = met, groupvars = 'value',
            narm = T) 
          names(tmp)[names(tmp) %in% met] <- 'mean'
          out[[met]] <- tmp

        }

        # combine metabolism categories 
        out <- melt(out, id.var = names(out[[1]]))
        names(out)[names(out) %in% 'L1'] <- 'var'
        out

      })

    # combine time categories
    agg_res <- melt(agg_res, id.var = names(agg_res[[1]]))
    names(agg_res)[names(agg_res) %in% 'L1'] <- 'cats'

    # separate var column into metabolis, obs/dtd
    agg_res$var[!grepl('dtd',agg_res$var)] <- paste0(
      agg_res$var[!grepl('dtd',agg_res$var)], '_obs')

    agg_res$sub_var <- gsub('^[A-Z,a-z]*_', '', agg_res$var)
    agg_res$var <- gsub('_[a-z]*$', '', agg_res$var)

    agg_res

  })

to_plo <- melt(agg_sum, id.var = names(agg_sum[[1]]))

# reassign factor labels
to_plo$value <- factor(to_plo$value, levels = c('01', '02', '03', '04',
  '05', '06', '07', '08', '09', '10', '11', '12', 'W', 'Sp', 'Su', 'F'), 
  labels = c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11',' 12',
    'W', 'Sp', 'Su', 'F')
  )
to_plo$var <- factor(to_plo$var, levels = c('Pg', 'Rt', 'NEM'), 
  labels = c('Pg', 'Rt', 'NEM'))
to_plo$sub_var <- factor(to_plo$sub_var, levels = c('dtd', 'obs'), 
  labels = c('Detided', 'Observed'))
to_plo$L1 <- gsub('_wtreg_[0-9]*.RData$', '', to_plo$L1)
to_plo$L1 <- factor(to_plo$L1, 
  levels = c('ELKVM', 'PDBBY', 'RKBMB', 'SAPDC'),
  labels = c('Elkhorn Slough', 'Padilla Bay', 'Rookery Bay', 
    'Sapelo Island')
  )

# convert metab data to g m^-2 d^-1
#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
to_plo[, c('mean', 'sd', 'se', 'ci')] <- 0.032 * to_plo[, c('mean', 'sd', 'se', 'ci')]

# remove may from Rookery Bay, lots of missing data
to_plo[to_plo$L1 %in% 'Rookery Bay'  & to_plo$value %in% '5', c('mean', 'sd', 'se', 'ci')] <- NA

ylab <- expression(paste('g ', O [2], ' ', m^-2, d^-1))
for(site in 3:4){

  sub_plo <- to_plo[to_plo$L1 %in% levels(to_plo$L1)[site], ]

  p <- ggplot(sub_plo, aes(x = factor(value), y = mean, group = sub_var,
      colour = sub_var)) +
    geom_line() +
    geom_point() + 
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), 
      width=.3) +
    facet_wrap(cats ~ var, scales = 'free') +
    theme_bw() +
    scale_y_continuous(ylab) + 
    scale_x_discrete(name = element_blank()) +
    theme(
      legend.position = 'top',
      legend.title = element_blank(), 
      plot.margin= unit(rep(0, 4), "lines"),
      axis.text = element_text(size = 8)
      ) +
    ggtitle(levels(to_plo$L1)[site])

  assign(paste0('p', site), p)

  }

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

# get common legend, remove from p1
mylegend <- g_legend(p3)
p3 <- p3 + theme(legend.position = 'none')
p4 <- p4 + theme(legend.position = 'none')

# final plot
grid.arrange(
  mylegend,   
  arrangeGrob(p3, p4, ncol = 1),
  ncol = 1, heights = c(1,10)
  )

@
\vfill
\clearpage

%%%%%%
% tables

\section{Tables}

% summary of simulation performance for detided and biological, sim parameters
<<dtd_perf1, cache = F, echo = F, eval = T, results = 'asis'>>=

# uses 'mod_perf' from, see readme
load('mod_perf.RData')

# calculate mean, se for dtd_cor and dtd_err
dtd_perf <- llply(names(mod_perf)[2:8], # note indexing
  .fun = function(x) {
    ddply(
      mod_perf,
      .variables = x,
      .fun = function(y) {
        c(
          min(y$cor),
          quantile(y$cor, 0.25),
          mean(y$cor),
          quantile(y$cor, 0.75),
          max(y$cor),
          min(y$err),
          quantile(y$err, 0.25),
          mean(y$err),
          quantile(y$err, 0.75),
          max(y$err)
        )
    })}
  )

# rename first column for combining, create column for parameter,  then combine
dtd_perf <- lapply(dtd_perf, 
  function(x) { 
    Parameter <- names(x)[1]
    names(x)[1] <- 'Level'
    x[,1] <- factor(x[,1])
    Parameter <- rep(Parameter, nrow(x))
    x <- data.frame(Parameter, x)
    x
    }
  )
dtd_perf <- do.call('rbind', dtd_perf)
names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Mean cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Mean err', 'High err', 'Max err')

# subset by DO time series conditions
to_tab <-dtd_perf[!dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]

# names for row categories in table
Parms <- c('$\\boldsymbol{DO_{die}}$', 
  '$\\boldsymbol{DO_{adv}}$',  
  '$\\boldsymbol{\\epsilon_{pro}}$', 
  '$\\boldsymbol{\\epsilon_{obs}}$'
  )
tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\\epsilon_{pro}$, $\\epsilon_{obs}$).  Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).'

# foot.val<-'\\footnotesize \\textit{Note:} test'

latex( 
  tab,
  file = '',
  where = 'h',
  rowlabel = 'Parameter',
#   insert.bottom = foot.val,
  caption = cap.val,
  caption.loc = 'top',
  rgroup = Parms,
  n.rgroup = rep(3, 4),
  cgroup = c('Correlation', 'RMSE'),
  n.cgroup = c(5, 5),
  rowname = rows,
  colheads=rep(c('Min', '25\\textsuperscript{th}', 'Mean', '75\\textsuperscript{th}', 'Max'),2),
  label = 'tab:dtd_perf1'
  )
@

% summary of simulation performance for detided and biological, window widths
<<dtd_perf2, cache = F, echo = F, eval = T, results = 'asis'>>=

# uses 'mod_perf' from 'dep_data.rnw'
load('mod_perf.RData')

# calculate mean, se for dtd_cor and dtd_err
dtd_perf <- llply(names(mod_perf)[1:8],
  .fun = function(x) {
    ddply(
      mod_perf,
      .variables = x,
      .fun = function(y) {
        c(
          min(y$cor),
          quantile(y$cor, 0.25),
          mean(y$cor),
          quantile(y$cor, 0.75),
          max(y$cor),
          min(y$err),
          quantile(y$err, 0.25),
          mean(y$err),
          quantile(y$err, 0.75),
          max(y$err)
        )
    })}
  )

# rename first column for combining, create column for parameter,  then combine
dtd_perf <- lapply(dtd_perf, 
  function(x) { 
    Parameter <- names(x)[1]
    names(x)[1] <- 'Level'
    x[,1] <- factor(x[,1])
    Parameter <- rep(Parameter, nrow(x))
    x <- data.frame(Parameter, x)
    x
    }
  )
dtd_perf <- do.call('rbind', dtd_perf)
names(dtd_perf) <- c('Parameter', 'Level', 'Min cor', 'Low cor', 'Mean cor', 'High cor', 'Max cor', 'Min err', 'Low err', 'Mean err', 'High err', 'Max err')

# subset by windows
to_tab <-dtd_perf[dtd_perf$Parameter %in% c('dec_time', 'hour', 'Tide'), ]

# names for row categories in table
Parms <- c('Days', 'Hours', 'Tide')

tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).  Values represent a combination of results from multiple simulations with the window value held constant for each row (e.g., row one is a summary of all simulations for which the half window width was one day).'

# foot.val<-'\\footnotesize \\textit{Note:} test'

latex( 
  tab,
  file = '',
  rowlabel = 'Window',
#   insert.bottom = foot.val,
  caption = cap.val,
  caption.loc = 'top',
  rgroup = Parms,
  n.rgroup = rep(3, 3),
  cgroup = c('Correlation', 'RMSE'),
  n.cgroup = c(5, 5),
  rowname = rows,
  colheads=rep(c('Min', '25\\textsuperscript{th}',  'Mean', '75\\textsuperscript{th}', 'Max'),2),
  label = 'tab:dtd_perf2'
  )
@

% descriptive table of case studies
<<case_att, cache = F, echo = F, eval = T, results = 'asis'>>=

# ######
# # table of case study characteristics
# # uses local files not in project....
# 
# cases <- c('ELKVM','PDBBY', 'RKBMB', 'SAPDC')
# 
# ##
# # get amps (m) of dominant tidal constituents
# 
# files <- list.files('M:/wq_models/SWMP/raw/rproc/proc5/', 
#   pattern = paste(cases, collapse = '|'), 
#   full.names = T)
# file_ls <- list()
# for(file in files){
#   load(file)
#   nm <- gsub('.RData', '', basename(file))
#   tmp <- get(nm)
#   mod <- oce::tidem(tmp$Depth, tmp$DateTimeStamp, 
#     constituents = c('P1', 'O1', 'M2', 'S2'))
#   const <- attr(mod, 'data')$amplitude[-1]
#   names(const) <- attr(mod, 'data')$name[-1]
#     
#   file_ls[[nm]] <- const
#   }
# tide_comps<- data.frame(do.call('rbind', file_ls))
# tide_comps$site <- rownames(tide_comps)
# 
# ##
# # get mean wq vals
# 
# files <- list.files('M:/wq_models/SWMP/raw/rproc/proc5/', 
#   pattern = paste(cases, collapse = '|'), 
#   full.names = T)
# file_ls <- list()
# for(file in files){
#   load(file)
#   nm <- gsub('.RData', '', basename(file))
#   cat(nm, '\t')
#   tmp <- get(nm)
#   file_ls[[nm]] <- tmp
#   }
# 
# wq_sums <- adply(
#   file_ls, 1, 
#   function(x) {
#     DO <- mean(x$DO_mgl, na.rm = T)
#     chl <- mean(x$CHLA_N, na.rm = T)
#     sal <- mean(x$Sal, na.rm = T)
#     temp <- mean(x$Temp, na.rm = T)
#     data.frame(DO, chl, sal, temp)
#     }
#   )
# names(wq_sums)[1] <- c('site')
# 
# ##
# # get metab summaries
# 
# # metab data
# # this is for observed data so, doesn't matter which index number
# load('met_ls.RData')
# 
# met_sum <- adply(matrix(cases, ncol = 1),
#   1,
#   .fun = function(x){
#     
#     met <- met_ls[[paste0(x, '_wtreg_1.RData')]][, c('Pg', 'Rt', 'NEM')]
#     colMeans(met, na.rm = T)
#      
#     }
#   )
# met_sum$X1 <- NULL
# met_sum$site <- cases
# 
# ##
# # combine data 
# # tide_comps, daily_do, met_sum
# 
# to_tab <- cbind(tide_comps, wq_sums, met_sum)
# to_tab <- to_tab[, !names(to_tab) %in% 'site']
# 
# case_tab <- to_tab
# save(case_tab, file = 'case_tab.RData')

load('case_tab.RData')

# convert metab data to g m^-2 d^-1
#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
tab <- case_tab
tab[, c('Pg', 'Rt', 'NEM')] <- 0.032 * tab[, c('Pg', 'Rt', 'NEM')]
tab <- form_fun(tab)
rows <- rownames(tab)

cap.val<-'Summary statistics of tidal component amplitudes (m), selected water quality parameters (\\ac{DO} mg L$^{-1}$, chlorophyll-a $\\mu$g L$^{-1}$, salinity psu, water temperature $^{\\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.  Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours), estimated from harmonic regressions of tidal height (\\texttt{oce} package in R, \\citealt{Foreman89}, \\citetalias{RDCT14}).  Water quality data are averages for the entire period of record (30 minute observations) for each site.  Metabolism estimates are means of daily integrated values.'

foot.val<-'\\footnotesize\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism'

latex( 
  tab,
  file = '',
  rowlabel = 'Site',
  insert.bottom = foot.val,
  caption = cap.val,
  caption.loc = 'top',
  cgroup = c('Tidal amplitude', 'Water quality', 'Metabolism\\textsuperscript{\\textit{a}}'),
  n.cgroup = c(4, 4, 3),
  rowname = rows,
  colheads=c('O1', 'P1', 'M2', 'S2', 'DO', 'Chl', 'Sal', 'Temp', 'Pg', 'Rt', 'NEM'),
  label = 'tab:case_att'
  )

@

% correlations with tide before/after wtreg
<<cor_res, cache = F, echo = F, eval = T, results = 'asis'>>=
# #####
# # table of DO/metab correlations with tide before after, detiding
# # note that tide in met_ls is daily average (day or night) of hourly tidal change
# 
# # metab and inst flux data
# load('met_ls.RData')
# # load('met_ls_inst.RData')
# load('case_grds.RData')
# 
# # go through each site for DO cors, use metab list for metab cors
# case_regs <- list.files('wtreg', '_wtreg_[0-9]*.RData')
# cor_res <- alply(matrix(case_regs),
#   1, 
#   .fun = function(x){
#   
#     # load wtreg data
#     load(paste0('wtreg/', x))
#     nm <- gsub('.RData', '', x)
#     dat_in <- get(nm)
#       
#     # DO obs v tide
#     do_obs <- with(dat_in, 
#       cor.test(DO_obs, Tide)
#       )
#     
#     # DO dtd v tide
#     do_dtd <- with(dat_in, 
#       cor.test(DO_nrm, Tide)
#       )
#     
#     # get tidal range for metabolic day/night periods from flux_in
#     # for correlation with daily integrated metab
#     tide_rngs <- ddply(dat_in, 
#       .variables = c('met.date'),
#       .fun = function(x){
#         
#         # mean tidal derivative for day hours
#         sunrise <- mean(diff(x[x$variable %in% 'sunrise', 'Tide'], 
#           na.rm = T))
#         
#         # mean tidal derivative for night hours
#         sunset <- mean(diff(x[x$variable %in% 'sunset', 'Tide'], 
#           na.rm = T))
#         if(sunrise == 'Inf') sunrise <- NA
#         if(sunset == 'Inf') sunset <- NA
#         
#         # mean tidal derivative for metabolic day
#         daytot <- mean(diff(x$Tide, na.rm = T))
#         
#         c(daytot, sunrise, sunset)
#         
#         }
#       )
#     names(tide_rngs) <- c('Date','daytot', 'sunrise', 'sunset')
#     
#     # get metab data from list
#     dat_in <- met_ls[[x]]
#     dat_in <- merge(dat_in, tide_rngs, by = 'Date', all.x = T)
#     
#     # as list for all metab correlations
#     # Pg values correlated with tidal range during sunlight hours
#     # Rt values correlated with tidal range during night hours
#     # NEM values correlated with metabolic daily tidal range
#     met_cor <- list(
#       
#       Pg_obs = with(dat_in, 
#         cor.test(Pg, sunrise)
#         ),
#     
#       Rt_obs = with(dat_in, 
#         cor.test(Rt, sunset)
#         ),
#     
#       NEM_obs = with(dat_in, 
#         cor.test(NEM, daytot)
#         ),
#     
#       Pg_dtd = with(dat_in, 
#         cor.test(Pg_dtd, sunrise)
#         ),
#     
#       Rt_dtd = with(dat_in, 
#         cor.test(Rt_dtd, sunset)
#         ),
#     
#       NEM_dtd = with(dat_in, 
#         cor.test(NEM_dtd, daytot)
#         )
#       
#       )
#     
#     # DO and metab corrs combined
#     all_ls <- c(do_obs = list(do_obs), do_dtd = list(do_dtd),
#       met_cor)
#     
#     # convert the stats for each wtreg to data frame
#     res_sum <- ldply(all_ls, 
#       function(x) with(x, c(estimate, p.value))
#       )
#     names(res_sum) <- c('var', 'cor', 'pval')
# 
#     res_sum
#     
#   })
# names(cor_res) <- case_regs
# 
# # melt and make separate columns for site and window comb value
# cor_res <- melt(cor_res, id.var = names(cor_res[[1]]))  
# cor_res$site <- gsub('_wtreg_[0-9]*.RData', '', cor_res$L1)
# cor_res$wins <- as.numeric(gsub('^.*_wtreg_|.RData', '', cor_res$L1))
# 
# # merge with case_grds
# case_grds$wins <- as.numeric(row.names(case_grds))
# cor_res <- merge(cor_res, case_grds, by = 'wins', all.x = T)
# 
# # create columns for variable (DO, flux, etc.) and sub variable (obs, dtd)
# cor_res$sub_var <- gsub('^.*_', '', cor_res$var)
# cor_res$var <- gsub('_.*$', '', cor_res$var)
# 
# save(cor_res, file = 'cor_res.RData')
# 
# to_plo <- cor_res
# to_plo$group_var <- paste(to_plo$Tide, to_plo$sub_var)
# to_plo_obs <- to_plo[to_plo$sub_var %in% 'obs', ]
# p1 <- ggplot(to_plo[to_plo$sub_var %in% 'dtd',], 
#     aes(x = dec_time, y = cor, colour = Tide, group = group_var)) +
#   geom_line() + 
#   geom_line(data = to_plo_obs, 
#     aes(x = dec_time, y = cor, group = group_var), 
#     colour = 'black', size = 1) +
#   geom_point(aes(pch = sub_var)) +
#   facet_grid(hour ~ site + var) +
#   ylim(c(-1, 1))

load('cor_res.RData')

# subset cor_res for window of 30 days, tidal prop 1
cor_res <- cor_res[with(cor_res, dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]), ]

# make wide form
to_tab <- dcast(cor_res, site + sub_var ~ var, value.var = 'cor')

# reorder rows, columns for plot
to_tab$sub_var <- factor(to_tab$sub_var, levels = c('obs', 'dtd'))
to_tab <- to_tab[with(to_tab, order(site, sub_var)), ]
to_tab <- to_tab[, c('site', 'sub_var', 'do', 'Pg', 'Rt', 'NEM')]

# p vals for each correlation
pvals <- dcast(cor_res, site + sub_var ~ var, value.var = 'pval')

# reorder rows, columns for plot
pvals$sub_var <- factor(pvals$sub_var, levels = c('obs', 'dtd'))
pvals <- pvals[with(pvals, order(site, sub_var)), ]
pvals <- pvals[, c('site', 'sub_var', 'do', 'Pg', 'Rt', 'NEM')]

# combine both tables
tab <- to_tab
rows <- as.character(tab$site)
tab <- form_fun(tab[, -c(1:2)])
pvals <- pvals[, -c(1:2)]
pdims <- dim(pvals)

pvals <- star_fun(c(as.matrix(pvals)))
tab <- c(as.matrix(tab))
tab <- paste0(tab, pvals)
tab <- matrix(tab, nrow = pdims[1], ncol = pdims[2])

cap.val<-'Correlations of tidal changes at each site with continuous \\ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (detided) detiding from weighted regression.  \\ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.'

foot.val<-'\\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\\\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism'

latex( 
  tab,
  file = '',
  rowlabel = 'Site',
  rgroup = unique(rows), 
  n.rgroup = rep(2, 4),
  insert.bottom = foot.val,
  caption = cap.val,
  colheads = c('DO', 'Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
  caption.loc = 'top',
  rowname = rep(c('Observed', 'Detided'), 4),
  label = 'tab:cor_res'
  )

@

% case study metabolism results, including perc anom
<<case_res, cache = F, echo = F, eval = T, results = 'asis'>>=
# # metab estimates for each wtreg
# load('met_ls.RData')
# 
# met_comp <- ldply(met_ls, function(x){
#   
#   # detided
#   anoms <- anoms.fun(x, pgvar = 'Pg_dtd', rtvar = 'Rt_dtd')
#   anoms <- data.frame(variable = 'Anom', anoms)
#   anoms <- melt(anoms, id.var = 'variable', variable.name = 'X1')[, c(2, 1, 3)]
#   subs <- na.omit(x[, c('Pg_dtd', 'Rt_dtd', 'NEM_dtd')])  
#   sums <- adply(t(subs),
#     1, 
#     function(y) data.frame(Avg = mean(y), se = sd(y)/sqrt(length(y)))
#     )
#   sums <- melt(sums, id.var = 'X1')
#   dtd <- rbind(anoms, sums)
#   dtd$X1 <- gsub('_dtd', '', dtd$X1)
#   dtd$Input <- 'Detided'
#   
#   # observed
#   anoms <- anoms.fun(x, pgvar = 'Pg', rtvar = 'Rt')
#   anoms <- data.frame(variable = 'Anom', anoms)
#   anoms <- melt(anoms, id.var = 'variable', variable.name = 'X1')[, c(2, 1, 3)]
#   subs <- na.omit(x[, c('Pg', 'Rt', 'NEM')])  
#   sums <- adply(t(subs),
#     1, 
#     function(y) data.frame(Avg = mean(y), se = sd(y)/sqrt(length(y)))
#     )
#   sums <- melt(sums, id.var = 'X1')
#   obs <- rbind(anoms, sums)
#   obs$Input <- 'Observed'  
#   
#   # output
#   out <- rbind(dtd, obs)
#   names(out) <- c('Metab', 'Metric', 'Value', 'Input')
#   
#   return(out)
#   
#   })
#
# 
# # subset metab estimates windows that were used
# load('case_grds.RData')
# sel_vec <- with(case_grds, which(dec_time == casewins[[1]] & hour == casewins[[2]] & Tide == casewins[[3]]))
# 
# met_comp <- met_comp[grep(sel_vec, met_comp$.id),]
# 
# save(met_comp, file = 'met_comp.RData')

load('met_comp.RData')

to_tab <- met_comp
names(to_tab)[1] <- 'Site'

# convert metab data to g m^-2 d^-1
#  1mmolO2 = 32 mg O2, 1000mg = 1g, multiply by 32/1000
sel_vec <- !to_tab$Metric %in% 'Anom'
to_tab$Value[sel_vec] <- 0.032 * to_tab$Value[sel_vec]

# make columns as factors for correct row, column order w/ dcast
to_tab$Input <- factor(to_tab$Input, levels = c('Observed', 'Detided'))
to_tab$Metab <- factor(to_tab$Metab, levels = c('Pg', 'Rt', 'NEM'))
to_tab$Metric <- factor(to_tab$Metric, levels = c('Avg', 'se', 'Anom'))

# dcast long to wide
to_tab <- dcast(to_tab, Site + Input ~ Metab + Metric, value.var = 'Value')

# remove extra text on site variable
to_tab$Site <- gsub('_wtreg_[0-9]*.RData$', '', to_tab$Site)

# names for row categories in table

tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \\ac{DO} time series before (observed) and after (detided) application of weighted regression.  Means and standard errors are based on daily integrated metabolism estimates.  Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.  Results are for weighted regression with half windows of six days, one hour around the time of observation for each day, and a tidal height proportion of one.'

foot.val<-'\\textsuperscript{\\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism' 
latex( 
  tab,
  file = '',
  rowlabel = 'Site',
  insert.bottom = foot.val,
  caption = cap.val,
  caption.loc = 'top',
  rgroup = unique(to_tab$Site),
  n.rgroup = rep(2, 4),
  cgroup = c('Pg\\textsuperscript{\\textit{a}}', 'Rt', 'NEM'),
  n.cgroup = c(3, 3, 2),
  rowname = rows,
  colheads=c(rep(c('Mean','Std. Err.', 'Anom'),2), c('Mean', 'Std. Err.')),
  label = 'tab:case_res'
  )
@
\clearpage

%%%%%%
% multimedia files, appendices

\raggedbottom
\raggedright
\setlength{\parindent}{0.5in}

\section{Multimedia} \label{multi}
\input{multimedia.tex}

\end{document}