### 2014-09-25 discussion.Rnw
final proof after MCM revisions

The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.  The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove variation in observed \ac{DO} from tidal change.  Further, aggregation of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location. These results suggest that previous estimates of annual means may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.  Additionally, variation of aggregated metabolism estimates were substantially reduced after filtering, suggesting greater confidence in interpreting estimates even if the mean values are similar.

Comparisons between filtered and biological \ac{DO} time series from the simulations indicated that weighted regression can reduce the effects of tidal variation for a range of characteristics of \ac{DO} time series.  An examination of scenarios that produced abnormal results can provide additional insight into factors that affect the performance of weighted regression.  For example, poor performance was observed when the observation uncertainty ($\epsilon_{obs}$) was high and both process uncertainty ($\epsilon_{pro}$) and tidal advection ($DO_{adv}$) were low.  These examples represent time series with excessive random variation, no auto-correlation, and no tidal influence.  Poor performance is expected because the weighted regression models a non-existent tidal signal in a very noisy \ac{DO} time series.  These results were observed even for time series with a large diel component of the biological \ac{DO} signal, suggesting that the model will produce random results in microtidal systems with high noise and no serial correlation.  From a practical perspective, weighted regression should not be applied to noisy time series if there is not sufficient evidence to suggest the variation is related to tidal changes.  Alternative approaches, such as the Kalman filter \citep{Harvey89,Batt12}, may be more appropriate if random variation is the primary source of uncertainty.  Similarly, results with perfect or near-perfect correlations between filtered and biological \ac{DO} time series were observed when observation uncertainty and tidal effects were not components of the simulated time series.  Although there is no need to apply weighted regression to time series with no apparent tidal influences, the results will not be incorrect.  We emphasize that the weighted regression should only be applied to time series for which specific conditions apply, as described in the recommendations below.

Correlations of metabolism estimates with tidal height changes after filtering were generally reduced, although trends were not always consistent.  However, correlations of net metabolism estimates were reduced in all cases.  An additional indication of the effectivenes of weighted regression was the reduction of anomalous metabolism estimates after filtering for all case studies.  Negative production and positive respiration estimates suggest assumptions of the open-water method are violated \citep{Needoba12}, although `normal' estimates (positive production and negative respiration) may still include a significant source of bias from physical advection by providing over-estimates of true values.  For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime such that inundation of salt marshes during the night following by draining with low tide during the day lead to inflated respiration values.   Synchrony between solar and tidal cycles is a critical concern for interpreting metabolism estimates, although a broader discussion regarding whether or not this represents an actual bias in metabolism from physical advection may be needed.

The weighted regression approach makes no assumptions as to the relationships between \ac{DO} and tidal variation over time.  Although the functional form of the model is a simple linear regression with only two explanatory variables (\cref{funform}), the moving window approach combined with the adaptive weighting scheme allows for quantification of complex tidal effects that may not be possible using alternative approaches.  A similer approach by \citet{Batt12} uses a Kalman filter to improve estimates of ecosystem metabolism in lakes.  The approach minimizes uncertainty in observed \ac{DO} using a filter that combines information about the data generation process and the manner in which the data are observed \citep{Harvey89}.  Although a similar approach could be used for estuaries, it may not be effective given that the effects of tidal advection are not related to process or observation uncertainty.  Additionally, results from the case studies illustrated the ability of the weighted regression approach to model changes over time in the relationships between tidal change and \ac{DO}.  Results for Padilla Bay and Rookery Bay suggested that filtering had the largest effect during the summer, whereas the results for cooler months were not significantly different from the observed.  The weighted regression method produced filtered time series that accommodated seasonal variation in \ac{DO} conditional on tidal height change, whereas moving window filters or standard regression techniques would likely not have characterized these dynamic relationships.