### 2015-01-28 comments.Rnw

Results from the simulations and case studies suggested that weighted regression can be a practical approach for filtering \ac{DO} time series.  However, application may only be appropriate under specific situations.  The case studies were chosen based on the relatively high percentage of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.  Despite these similarites, filtering had variable effects on metabolism estimates.  The results for Elkhorn Slough and Padilla Bay are of particular concern given that annual mean estimates were substantially different compared to those from the observed \ac{DO} time series.  Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, collinearity between tidal change and solar cycling likely contributed to the relative change in mean metabolism before and after filtering.  Evaluating collinearity, as in \cref{fig:move_corrs}, is an important diagnostic for characterizing the expected performance of weighted regression and choosing appropriate window widths.  These plots should indicate whether weighted regression is appropriate for the whole time series (e.g., Sapelo Island) or only for specific periods when collinearity is low (e.g., April and October at Elkhorn Slough).

\ac{DO} time series may include components of metabolism related to true biological effects linked to the tide, in addition to diurnal cycling and effects of water movement.  For example, increased respiration may be observed during ebb tides as a result of increased enrichment of substrates \citep{Sasaki09}.  In such cases, variation in the metabolic signal is correlated and directly related to the tide but is not related to physical movement of water masses. Accordingly, the model should not be used without critical evaluation of site-level characteristics that suggest physical movement of water masses is a primary source of noise in a \ac{DO} signal. Locations where tidal influences are large and the water is relatively deep (as in the Vierra Mouth station at Elkhorn Slough) are clear examples where physical movement is likely the dominant source of variation in the \ac{DO} signal.  Application to stations where this is less clear should be carefully evaluated, particularly with respect to changing window widths on the results.  In general, the results should be evaluated using multiple performance metrics as above. The optimal window widths are likely those that reduce anomalous metabolism estimates, minimize correlations between observed \ac{DO} and tide height, and reduce clear visual patterns of tide change on \ac{DO}.  In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.  Such an approach, combined with further validation, will support informed scientific evaluations by providing more accurate estimates of ecosystem metabolism.