### 2015-02-09 comments.Rnw

Results from the analysis of simulated data and case studies suggest that weighted regression is a useful approach for improving estimates of ecosystem metabolism by reducing the effects of tidal advection on \ac{DO} time series.  Case study sites were selected that, when ecosystem metabolism was calculated using standard open water methods, had a large number of observations  that were not interpretable (i.e., `anomalous`).  In other words, the advection effects on \ac{DO} time series at these sites are very large.  Our weighted regression method dramatically reduced advection effects at some of these sites, generating useful estimates of metabolism where otherwise open water metabolism estimates would be subject to significant errors.  Further analysis of the method illustrated the attributes of sites where it should be most useful and conversely, attributes of sites where problems may be encountered using the method.  The weighted regression method will work best at sites that have a medium or large effect of advection on \ac{DO}, but tidal changes are not correlated with the solar cycle for more than a few days.  At sites with a dominant semidiurnal tide, such as at Sapelo Island, Georgia, the method can be expected to dramatically improve analysis using open water metabolism.  Since this tidal regime occurs throughout the Atlantic coast of North America, Europe and Africa, as well as many other areas of the global coastline, this method should be broadly applicable, even if not universally applicable.  Where tidal effects on \ac{DO} time series are small, the method has the potential to reduce advection effects further, although the potential for improvement is smaller than at sites with more significant advection effects.  Where tides are large and strongly correlated with the solar cycle on seasonal time scales, the traditional open water metabolism method may be subject to very large errors.  This was observed at the Elkhorn Slough and Padilla Bay sites, which have mixed semidiurnal tides.  Tides with this pattern are characteristic of the Pacific coastline of North America and portions of east Asia, among others.  Unfortunately, the weighted regression method could not resolve which \ac{DO} changes resulted from advection and which resulted from metabolism at these sites because the patterns are excessively collinear.  As we implemented the method, the regression attributed most \ac{DO} variation at these sites to the tide, resulting in estimates of metabolism that are very small.  In these cases, the computed metabolism estimates after weighted regression are different from the original open water metabolism computations, but equally incorrect.

\ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.  For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.  Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.  The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.  We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.

As a final note, it is worth mentioning that these computations are complex relative to the basic open water metabolism calculations which can be implemented by hand or in a spreadsheet.  However, because we implemented the analysis in R, the freely-available programming environment for statistical analysis, scientific computing and graphics \citepalias{RDCT14}, many users may be able to replicate the approach.  As the original open-water metabolism computations have already been implemented in R and shared widely, the additional development of the method presented here could also be applied by using our code.