### 2015-05-18 multimedia.tex

The supporting information for this manuscript includes a graphical illustration of the weighting scheme described in the material and procedures section (\href{http://spark.rstudio.com/beckmw/weights_widget}{http://spark.rstudio.com/beckmw/weights\_widget}), results for each simulation (\href{http://spark.rstudio.com/beckmw/detiding_sims}{http://spark.rstudio.com/beckmw/detiding\_sims}), and results for each case study (\href{http://spark.rstudio.com/beckmw/detiding_cases}{http://spark.rstudio.com/beckmw/detiding\_cases}).  Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.

A simple R package with a sample dataset and code to implement weighted regression is also available on GitHub.  Functions are also availabe to estimate ecosystem metabolism.  See the README file on the web page for download instructions and examples: \href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}