### 2015-05-24 cover_letter.tex
updated cover letter for resubmission

\documentclass[a4paper,12pt]{article} 
\usepackage{times}
\usepackage[top=1in,bottom=1in,left=1in,right=1in,nohead]{geometry}
\usepackage{graphics}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}

\begin{document}
\renewcommand{\rmdefault}{ptm}
\pagestyle{empty}

\setlength{\parindent}{0mm} 
\setlength{\parskip}{5mm}

\begin{flushright}
\today
\end{flushright}

\emph{To:}\newline
Dr. Paul F. Kemp\newline
Editor-in-Chief\newline
Limnology and Oceanography: Methods, ASLO

\emph{From:}\newline
Dr. Marcus W. Beck\newline
ORISE Research Participation Program\newline
US Environmental Protection Agency\newline
beck.marcus@epa.gov, 850-934-2480\vspace{14.5pt}

\raggedright

Enclosed please find our revised manuscript entitled `Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series'.  We sincerely thank the reviewers and associate editor for providing detailed and insightful comments on our initial draft.  We have made every effort to address these concerns and are hopeful that the revised manuscript is an improvement over the original.  In particular, we have made efforts to more carefully describe appropriate situations for applying the weighted regression approach, additional details describing our methods for estimating open-water metabolism, and more justification for the use of tidal height as a proxy for advection.  We have also created a simple \href{https://github.com/fawda123/WtRegDO}{R package} to accompany the manuscript in response to the request for making the code available.  Our response to the reviewer comments is included as an attachment to this re-submission.

All authors have contributed equally to the content of the revised manuscript and have approved the final product.  The manuscript contains only original data and the authors agree to pay publication charges if accepted.  We greatly appreciate the opportunity to publish our work in Limnology and Oceanography: Methods.\vspace{0.5in}

\hspace{4.5in}Respectfully,

\hspace{4.5in}Marcus W. Beck

\end{document}